A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-02 latex/main.tex
Updates from Overleaf

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
We reconstructed
a more comprehensive mutational landscape of SARS-CoV-2
by mapping deletions onto the ARG using post-hoc parsimony.
We identified major deletion events, for example,
the deletions inherited by at least 10,000 samples
shown in Table~\ref{tab:major_dels}.
% This just says that the deletions are found in sequences of VOCs.
All these deletions are known to occur in major VOCs 
(\url{https://virological.org/t/preliminary-genomic-characterisation-of-an-emergent-sars-cov-2-lineage-in-the-uk-defined-by-a-novel-set-of-spike-mutations/563};
\url{https://github.com/cov-lineages/pango-designation/issues/361};
\cite{Stern2021,Tegally2022,Li2023indels}),
except the 4-base deletion and two 5-base deletions in ORF1ab,
which are likely artefacts of
deletion remapping or sequence misalignment (see below).
Notably, eight of these deletions occur
in the N-terminal domain of the S1 unit in Spike,
a known hotspot for recurrent deletions \cite{McCarthy2021}
(also see Table~\ref{tab:rec_dels}).
In the ARG, we found that most of the major deletion events
are associated with the origins of major VOCs
(Table~\ref{tab:major_dels}).
They have been linked previously with the emergence of these VOCs
(\url{https://virological.org/t/preliminary-genomic-characterisation-of-an-emergent-sars-cov-2-lineage-in-the-uk-defined-by-a-novel-set-of-spike-mutations/563};
\url{https://github.com/cov-lineages/pango-designation/issues/361};
\cite{Stern2021,Tegally2022}).

@@

\begin{table}
\centering
\caption{
Major deletion events in the ARG.
These events are inherited by
at least 10,000 samples in the ARG.
Count is the number of samples inheriting a deletion.
The Pango lineage assigned to the node
above which an event occurs is shown.
Abbreviations: NSP, non-structural protein;
ORF, open reading frame;
S, Spike;
S1-NTD, N-terminal domain in Spike S1 subunit.
}
\begin{tabular}{rrcrcrrc}
\toprule
Start & Length & Region & Node & Pango lineage & Count & Frequency (\%) & Ref\\
\midrule
28271 & 1 & non-coding & 1436808 & B.1 & 1118363 & 45.0561 &  \\
28248 & 6 & ORF8 & 220186 & B.1.617.2 (Delta) & 1114669 & 44.9073 & B \\
22029 & 6 & S / S1-NTD & 200039 & B.1.617.2 (Delta) & 1112504 & 44.8200 & B \\
11288 & 4 & ORF1ab / NSP6 & 1436802 & B.1.1.529 (Omicron) & 871782 & 35.1220 &  \\
28362 & 9 & N & 1436802 & B.1.1.529 (Omicron) & 871774 & 35.1216 & C \\
21633 & 9 & S / S1-NTD & 822854 & BA.2 (Omicron) & 532072 & 21.4359 & D \\
11292 & 5 & ORF1ab / NSP6 & 822854 & BA.2 (Omicron) & 531894 & 21.4287 &  \\
6513 & 3 & ORF1ab / NSP3 & 851246 & BA.1 (Omicron) & 340757 & 13.7283 & D \\
11283 & 5 & ORF1ab / NSP6 & 851246 & BA.1 (Omicron) & 339867 & 13.6924 &  \\
21765 & 6 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339681 & 13.6849 & D \\
21987 & 9 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339651 & 13.6837 & D \\
22194 & 3 & S / S1-NTD & 851246 & BA.1 (Omicron) & 312706 & 12.5982 & D \\
28271 & 1 & non-coding & 86456 & B.1.1.7 (Alpha) & 290446 & 11.7014 &  \\
11288 & 9 & ORF1ab / NSP6 & 86456 & B.1.1.7 (Alpha) & 290408 & 11.6998 & A \\
21991 & 3 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 288831 & 11.6363 & A \\
21765 & 6 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 290438 & 11.7010 & A \\
21765 & 6 & S / S1-NTD & 1265302 & BA.4 (Omicron) & 190425 & 7.6718 & E \\
686 & 9 & ORF1ab / NSP1 & 2698748 & BA.4 (Omicron) & 23330 & 0.9399 &  \\
\bottomrule
\end{tabular}
\end{table}

>>>>

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
We reconstructed
a more comprehensive mutational landscape of SARS-CoV-2
by mapping deletions onto the ARG using post-hoc parsimony.
We identified major deletion events, for example,
the deletions inherited by at least 10,000 samples
shown in Table~\ref{tab:major_dels}.
% This just says that the deletions are found in sequences of VOCs.
All these deletions are known to occur in major VOCs
\cite{RambautAlpha2020,Stern2021,RambautOmicron2021,
Tegally2022,Li2023indels},
except the 4-base deletion and two 5-base deletions in ORF1ab,
which are likely artefacts of
deletion remapping or sequence misalignment (see below).
Notably, eight of these deletions occur
in the N-terminal domain of the S1 unit in Spike,
a known hotspot for recurrent deletions \cite{McCarthy2021}
(also see Table~\ref{tab:rec_dels}).
In the ARG, we found that most of the major deletion events
are associated with the origins of major VOCs
(Table~\ref{tab:major_dels}), and
they have been linked previously with the emergence of these VOCs
\cite{RambautAlpha2020,Stern2021,RambautOmicron2021,Tegally2022}.

@@

\begin{table}
\centering
\caption{
Major deletion events in the ARG.
These events are inherited by
at least 10,000 samples in the ARG.
Count is the number of samples inheriting a deletion.
The Pango lineage assigned to the node
above which an event occurs is shown.
Abbreviations: NSP, non-structural protein;
ORF, open reading frame;
S, Spike;
S1-NTD, N-terminal domain in Spike S1 subunit.
}
\label{tab:major_dels}
\begin{tabular}{rrcrcrrc}
\toprule
Start & Length & Region & Node & Pango lineage & Count & Frequency (\%) & Ref\\
\midrule
28271 & 1 & non-coding & 1436808 & B.1 & 1118363 & 45.0561 &  \\
28248 & 6 & ORF8 & 220186 & B.1.617.2 (Delta) & 1114669 & 44.9073 & \cite{Stern2021} \\
22029 & 6 & S / S1-NTD & 200039 & B.1.617.2 (Delta) & 1112504 & 44.8200 & \cite{Stern2021} \\
11288 & 4 & ORF1ab / NSP6 & 1436802 & B.1.1.529 (Omicron) & 871782 & 35.1220 &  \\
28362 & 9 & N & 1436802 & B.1.1.529 (Omicron) & 871774 & 35.1216 & C \\
21633 & 9 & S / S1-NTD & 822854 & BA.2 (Omicron) & 532072 & 21.4359 & D \\
11292 & 5 & ORF1ab / NSP6 & 822854 & BA.2 (Omicron) & 531894 & 21.4287 &  \\
6513 & 3 & ORF1ab / NSP3 & 851246 & BA.1 (Omicron) & 340757 & 13.7283 & D \\
11283 & 5 & ORF1ab / NSP6 & 851246 & BA.1 (Omicron) & 339867 & 13.6924 &  \\
21765 & 6 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339681 & 13.6849 & D \\
21987 & 9 & S / S1-NTD & 851246 & BA.1 (Omicron) & 339651 & 13.6837 & D \\
22194 & 3 & S / S1-NTD & 851246 & BA.1 (Omicron) & 312706 & 12.5982 & D \\
28271 & 1 & non-coding & 86456 & B.1.1.7 (Alpha) & 290446 & 11.7014 &  \\
11288 & 9 & ORF1ab / NSP6 & 86456 & B.1.1.7 (Alpha) & 290408 & 11.6998 & A \\
21991 & 3 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 288831 & 11.6363 & A \\
21765 & 6 & S / S1-NTD & 86456 & B.1.1.7 (Alpha) & 290438 & 11.7010 & A \\
21765 & 6 & S / S1-NTD & 1265302 & BA.4 (Omicron) & 190425 & 7.6718 & E \\
686 & 9 & ORF1ab / NSP1 & 2698748 & BA.4 (Omicron) & 23330 & 0.9399 &  \\
\bottomrule
\end{tabular}
\end{table}
